Embedding Dimension of Polyhedral Losses

### 2020-01-29 colt-20/embed-dim.tex
edits to beginning of geometric abstain proof

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}\label{sec:examples}
\subsection{Abstain property, $\alpha \leq 1/2$}\label{subsec:example-abstain}
One classification-like problem that is of particular interest is the abstain property, elicited by the loss $\ellabs{1/2}$ given in Equation~\ref{eq:abstain}.
The property $\abstain{\alpha}$ can be verified:
\begin{equation}\label{eq:abstain-prop}
     \abstain{\alpha} = \begin{cases}
     \argmax_{y \in \Y} p(y) & \max_y p(y) \geq 1 - \alpha\\ 
     \bot & \text{otherwise}
     \end{cases}~.~
\end{equation}

@@

With our $1$-dimensional characterization, we already observe a tighter lower bound.
\begin{proposition}
	For $n \geq 3$ and $\alpha \leq 1/2$, we have $\elicembed(\abstain{1/2}) \geq 2$.
\end{proposition}
\begin{proof}
	Consider the intersection graph of $\gamma := \abstain{\alpha}$: the node associated with $\gamma_\bot$ has $n$ edges, and since we assume $n \geq 3$, we fail to observe the requirement that every node of the intersection graph has degree at most $2$.
\end{proof}

\begin{proposition}
	$\gamma := \abstain{1/2}$ with $n=5$ is not $2$-embeddable.
\end{proposition}
\begin{proof}
  We use the results from Section~\ref{sec:d-dim} and show that the optimality condition is not met for the level set $\gamma_\bot$.
  First, we represent the cell $\gamma_\bot = \{p \in \simplex : Bp \geq \vec 0\} = \conv(\{(1/2, 1/2, 0,0,0), \ldots, (0,0,0,1/2, 1/2)\})$, where $B = \ones \ones^T - 2I_5 \in \reals^{5 \times 5}$ has entry $-1$ on the diagonal, and entry $1$ everywhere else.

  We make an argument on the permissible angles between normals $v_1, \ldots, v_5$ in $\reals^2$ and show there is no feasible solution to the QFP in Definition~\ref{def:qfp}.
  In particular, we argue that for any cone strictly $ < 180^\circ$, we can have no more than two vectors $v_i$ and $v_j$; otherwise we yield a contradiction.

  Without loss of generality, we let $v_1 = (0,1)$, and have $v_2$ with $\|v_2\|^2 \leq \|v_1\|^2$ and the angle $\theta_{12} < 180^\circ$ the angle between $v_1$ and $v_2$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}\label{sec:examples}
\bo{I suggest we don't use the notation $\elicembed$ at all and just write out $d$-embeddable every time. But don't feel strongly.}
\subsection{Abstain property, $\alpha \leq 1/2$}\label{subsec:example-abstain}
One classification-like problem that is of particular interest is the abstain property, elicited by the loss $\ellabs{1/2}$ given in Equation~\ref{eq:abstain}.
The property $\abstain{\alpha}$ can be verified:
\begin{equation}\label{eq:abstain-prop}
     \abstain{\alpha} = \begin{cases}
     \argmax_{y \in \Y} p(y) & \max_y p(y) \geq 1 - \alpha\\ 
     \bot & \text{otherwise}
     \end{cases}~.~
\end{equation}

@@

With our $1$-dimensional characterization, we already observe a tighter lower bound.
\begin{proposition}
	For $n \geq 3$ and $\alpha \leq 1/2$, the abstain loss $\ell_{\alpha}$ is not $1$-embeddable. \bo{In fact, for any $\alpha < 1$...}
\end{proposition}
\begin{proof}
	Consider the intersection graph of $\gamma := \abstain{\alpha}$: the node associated with $\gamma_\bot$ has $n$ edges, and since we assume $n \geq 3$, we fail to observe the requirement that every node of the intersection graph has degree at most $2$.
\end{proof}

\begin{lemma} \label{lemma:abstain-v-h}
  For the abstain loss $\ell_{1/2}$, the level set of abstain satisfies $\gamma_{\bot} = \conv\{(\delta_y + \delta_{y'})/2 : y,y' \in \Y, y\neq y'\} = \{p : Bp \geq \vec{0}\}$ where $\delta_y$ puts probability one on $y$ and $B = \ones \ones^{\intercal} - 2I \in \reals^{5 \times 5}$, i.e. has entries $-1$ on the diagonal and $1$ everywhere else.
\end{lemma}
\begin{proof}
  Recall that $\gamma_{\bot}$ is the set of distributions $p$ with $\max_y p(y) = 1/2$.
  First, note that each distribution of the form $(1/2, 1/2, 0, 0, 0)$ and so on is in $\gamma_{\bot}$.
  Meanwhile, every such $p$ can be written as a convex combination of these corners.
  Second, note that if $p \in \gamma_{\bot}$, then $p(y) \leq 1/2$ for all $y \in \Y$.
  These constraints can be rewritten as $\inprod{p}{b} \geq 0$ where $b_y = -1$ and $b_{y'} = 1$ for all $y' \neq y$, literally requiring $p(y) \leq \sum_{y' \neq y} p(y')$.\bo{Check notation $p_y$ versus $p(y)$.}
\end{proof}

\begin{proposition}
	The abstain loss $\ell_{1/2}$ with $n=5$ is not $2$-embeddable.
\end{proposition}
\begin{proof}[Geometric proof]
  We apply Corollary \ref{cor:d-embeddable-char}: We consider the level set $\gamma_{\bot}$ (the distributions whose modal mass is at most $1/2$).
  We show that the quadratic program (Definition \ref{def:qfp}) is not feasible for this polytope with dimension $d=2$, which implies by Corollary \ref{cor:d-embeddable-char} that $\ell$ is not $2$-embeddable.

  We obtain the vertex and halfspace representations of $\gamma_{\bot}$ from Lemma \ref{lemma:abstain-v-h}.
  We proceed by narrowing the space of feasible normals $v_1, \ldots, v_5$, until we reach a contradiction.
  %We make an argument on the permissible angles between normals $v_1, \ldots, v_5$ in $\reals^2$ and show there is no feasible solution to the QFP in Definition~\ref{def:qfp}.
  In particular, we argue that in any cone strictly $ < 180^\circ$, we can have no more than two vectors $v_i$ and $v_j$; otherwise we yield a contradiction.

  We index the vertices and witnesses as follows: For $y < y'$, we write $p^{yy'} = (\delta_y + \delta_{y'})/2$ and $X^{yy'}$ for the corresponding matrix variable, with columns (witness points) $x^{yy'}_{y''}$.
  In particular, the vertex constraints can all be rearranged to yield the following: $\{x^{yy'}_y = -x^{yy'}_{y'} ~ (\forall y < y')\}$.
  Our proof will ignore all other points, i.e. $x^{yy'}_{y''}$ for $y'' \neq y,y'$, which appear only with coefficient zero in the vertex constraints.

  Without loss of generality, we let $v_1 = (0,1)$, and have $v_2$ with $\|v_2\|^2 \leq \|v_1\|^2$ and the angle $\theta_{12} < 180^\circ$ the angle between $v_1$ and $v_2$. \bo{Need to justify why choice of $v_1$ is WLOG.}
